---
title: "milestone 2"
author: "Huailin Tang"
date: "3/10/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("ggplot2")
```

# vaccine
```{r}
# data source: https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh
vaccine <- read.csv("COVID-19_Vaccinations_in_the_United_States_County.csv")
```

```{r}
head(vaccine)
```

```{r}
vaccine$Date = as.Date(vaccine$Date, format = "%m/%d/%Y")
vaccination_new <- vaccine %>% 
  filter(Date >= "2022/03/01" & Date <= "2022/03/31") %>% #only use march data
  filter(Recip_State %in% c("ND","SD","NE","KS","MO","IA","MN","WI","MI","IL","IN","OH")) %>% #filter Midwest
  select(Date, Recip_County, Recip_State, Administered_Dose1_Pop_Pct, Series_Complete_Pop_Pct, Metro_status) %>% #filter key variable
  rename(county = Recip_County, state = Recip_State, dose_1_pct = Administered_Dose1_Pop_Pct, fully_pct = Series_Complete_Pop_Pct, metro = Metro_status)
```

```{r}
head(vaccination_new)
```

```{r}
vaccination_new %>% 
  ggplot() +
  geom_point(aes(x=Date, y=dose_1_pct)) 

vaccination_new %>% 
  ggplot() +
  geom_point(aes(x=Date, y=fully_pct)) 

vaccination_new %>% 
  ggplot() +
  geom_point(aes(x=state, y=dose_1_pct))

vaccination_new %>% 
  ggplot() +
  geom_point(aes(x=state, y=fully_pct))

vaccination_new %>% 
  ggplot() +
  geom_point(aes(x=county, y=dose_1_pct))

vaccination_new %>% 
  ggplot() +
  geom_point(aes(x=county, y=fully_pct))
```

From the visualization below, we see there are zero dose 1 percentage and fully vaccination percentage. These zero might from the same place. 
After we look at the data, we see there are many "Unknown County", we get rid of them. It is likely these "Unknown County" is caused by data collection errors. 
There is also NA in the dataset, we replace the NA with the average of vaccination rate (dose 1 percentage or fully vaccination percentage) of counties in the same state, in the same date. 
There are still some zero left in dose 1 percentage. More specifically, all dose 1 percentage in NE from 03/01 to 03/15 are 0. This might be caused by NE does not report dose 1 percentage to CDC at this time period. We will not include NE in the graph for dose 1 percentage.
Some Michigan counties: Montcalm county from 03/02 to 03/15, Newayago county from 03/01 to 03/15, and Van Buren county from 03/05 to 03/15 all have 0 dose 1 percentage. We replace them with nearest non-zero first dose percentage.

```{r}
vaccination_new <- vaccination_new %>%
  filter(county != "Unknown County")

#replace NA in dose_1_pct with average
for(i in 1:nrow(vaccination_new)){
  if (is.na(vaccination_new[i,"dose_1_pct"]) == TRUE){
    dose_1_mean = vaccination_new %>% 
      filter(Date == vaccination_new[i, "Date"], 
             state == vaccination_new[i, "state"]) %>% 
      summarise(mean = mean(vaccination_new[,"dose_1_pct"], na.rm = TRUE)) %>% 
      pull(unique(mean))
    vaccination_new[i,"dose_1_pct"] = dose_1_mean
  }
}

#replace NA in fully_pct with average
for(i in 1:nrow(vaccination_new)){
  if (is.na(vaccination_new[i,"fully_pct"]) == TRUE){
    fully_mean = vaccination_new %>% 
      filter(Date == vaccination_new[i, "Date"], 
             state == vaccination_new[i, "state"]) %>% 
      summarise(mean = mean(vaccination_new[,"fully_pct"], na.rm = TRUE)) %>% 
      pull(unique(mean))
    vaccination_new[i,"fully_pct"] = fully_mean
  }
}

#replace zero value in three county
for(i in 1:nrow(vaccination_new)){
  if (vaccination_new[i, "county"] == "Van Buren County" & 
      vaccination_new[i, "state"] == "MI" &
      vaccination_new[i, "Date"] >= "2022-03-05" & 
      vaccination_new[i, "Date"] <= "2022-03-15"){
    vaccination_new[i, "dose_1_pct"] = 56.7 #replace with dose 1 percentage on 03/04
  }
  if (vaccination_new[i, "county"] == "Montcalm County" & 
      vaccination_new[i, "state"] == "MI" &
      vaccination_new[i, "Date"] >= "2022-03-02" & 
      vaccination_new[i, "Date"] <= "2022-03-15"){
    vaccination_new[i, "dose_1_pct"] = 42.1 #replace with dose 1 percentage on 03/01
  }
  if (vaccination_new[i, "county"] == "Newaygo County" & 
      vaccination_new[i, "state"] == "MI" &
      vaccination_new[i, "Date"] >= "2022-03-01" & 
      vaccination_new[i, "Date"] <= "2022-03-15"){
    vaccination_new[i, "dose_1_pct"] = 54.49181 #replace with dose 1 percentage on 03/16
  }
}
```

```{r}
write.csv(vaccination_new,"vaccination_new.csv", row.names = FALSE)
```

# COVID-19 transmission
```{r}
# data source: https://data.cdc.gov/Public-Health-Surveillance/United-States-COVID-19-County-Level-of-Community-T/nra9-vzzn
transmission <- read.csv("United_States_COVID-19_County_Level_of_Community_Transmission_Historical_Changes.csv")
```

```{r}
head(transmission)
```

```{r}
transmission$date = as.Date(transmission$date, format = "%m/%d/%Y")
transmission_new <- transmission %>% 
  filter(date >= "2022/03/01" & date <= "2022/03/31") %>% 
  filter(state_name %in% c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin")) %>% 
  select(state_name, county_name, date, percent_test_results_reported_positive_last_7_days, community_transmission_level) %>% 
  rename(positive_pct = percent_test_results_reported_positive_last_7_days, transmission_level = community_transmission_level)
```

```{r}
transmission_new %>% 
  ggplot() +
  geom_point(aes(x=state_name, y=positive_pct))

transmission_new %>% 
  filter(state_name == "Iowa") #NA for all positive_pct

transmission_new %>% 
  ggplot() +
  geom_point(aes(x=county_name, y=positive_pct))

transmission_new %>% 
  ggplot() +
  geom_point(aes(x=date, y=positive_pct))
```

I examine the each variable in the reorganized dataset. 
First problem: Iowa don't have Covid-19 test results. We exclude Iowa in the plot.
Second problem: some rows have 0 in positive_pct. This might be caused by low testing number in the county. 
Third problem: some rows have NA. A possible solution is replacing the NA with the average of positive_pct of counties with same transmission level, in the same state and in the same date.

Fix the dataset based on the problems and solutions above.

```{r}
transmission$date = as.Date(transmission$date, format = "%m/%d/%Y")
transmission_new <- transmission %>% 
  filter(date >= "2022-03-01" & date <="2022-03-31") %>% # change the date
  filter(state_name %in% c("Illinois", "Indiana", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin")) %>% 
  select(state_name, county_name, date, percent_test_results_reported_positive_last_7_days, community_transmission_level) %>% 
  rename(positive_pct = percent_test_results_reported_positive_last_7_days, transmission_level = community_transmission_level)

# replace NA with mean
for(i in 1:nrow(transmission_new)){
  if (is.na(transmission_new[i,"positive_pct"]) == TRUE){
    transmission_mean = transmission_new %>% 
      filter(date == transmission_new[i, "date"], 
             state_name == transmission_new[i, "state_name"],
             transmission_level == transmission_new[i, "transmission_level"]) %>% 
      summarise(mean = mean(transmission_new[,"positive_pct"], na.rm = TRUE)) %>% 
      pull(unique(mean))
    transmission_new[i,"positive_pct"] = transmission_mean
  }
}
```

```{r}
write.csv(transmission_new,"transmission_new.csv", row.names = FALSE)
```


